---
title: "MSME_loans"
author: "Osasere"
date: "2024-01-09"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MSME LOANS

The government have decided to give loans to Micro, small and medium enterprises in the state.

```{r cars}
summary(cars)
```

#there are some columns with inappropraite column names so we are going to standardize the column names
df <- df %>%
  rename(sen_zone = `Sen. Zone`, amount_granted = `Amount Granted`)
View(df)
df <- df %>%
  rename(
    principal_balance = `Principal Balance`, date_granted = `Date Granted`,
    expiry_date = `Expiry Date`, balance = Balance,
    amount_paid = `Paid amount`, total_amount = `Total amount to be paid`)
view

#dropping unwanted columns in the data frame
library(dplyr)

df <- df %>%
  select(-(`Principal paid`), -('Interest paid'))

df <- df %>%
  select(-(`Principal past due`), -('Interest past due'))

#summary statistics for descriptive statistics
summary(df)

attach(df)

#to understand the number of males and females that benefited from this project

table<-table(Sex)

ggplot(df, aes(x = Sex, fill = Sex)) +
  geom_bar(stat = "count", show.legend = FALSE) +  
  geom_text(stat = "count", aes(label = ..count..), position = position_stack(vjust = 1.02)) +  
  scale_fill_manual(values = c("red", "blue"), name = "Sex") + 
  labs(title = "Number of Beneficiaries per Gender", x = "Sex", y = "Number of Beneficiaries")

#Boxplot to show the distribution of amount granted per sex in this data
ggplot(df, aes(x = Sex, y = amount_granted, fill = Sex)) +
  geom_boxplot() +
  scale_fill_manual(values = c("red", "blue"), name = "Sex") +  
  labs(title = "Amount Granted per Sex", x = "Sex", y = "Amount Granted") 

#Boxplot to show the distribution of amount granted per Senatorial zones 
ggplot(df, aes(x = sen_zone, y = amount_granted, fill = sen_zone)) +
  geom_boxplot() +
  scale_fill_manual(values = c("red", "blue","yellow"), name = "sen_zone") +  
  labs(title = "Amount Granted per Senatorial Zone", x = "sen_zone", y = "Amount Granted") 

#to specifically show the average amount granted by sex
mean_amount_by_sex <- df %>%
  group_by(Sex) %>%
  summarize(mean_amount = mean(amount_granted, na.rm = TRUE))

print(mean_amount_by_sex)

#Barplot to visualise the difference in mean of the amount granted for different sexes
ggplot(df, aes(x = Sex, y = amount_granted, fill = Sex)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  scale_fill_manual(values = c("red", "blue"), name = "Sex") + 
  labs(title = "Mean Amount Granted by Sex", x = "Sex", y = "Mean Amount Granted") 

#Barplot to visualise the difference in mean of the amount granted for senatorial zones
ggplot(df, aes(x = sen_zone, y = amount_granted, fill = sen_zone)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  scale_fill_manual(values = c("red", "blue", "yellow"), name = "Sen_zone") + 
  labs(title = "Mean Amount Granted by Senatorial Zone", x = "Senatorial Zones", y = "Mean Amount Granted") 

#Barplot to visualise the sum of the amount granted by sexes
ggplot(df, aes(x = Sex, y = amount_granted, fill = Sex)) +
  geom_bar(stat = "summary", fun = "sum", position = "dodge") +
  scale_fill_manual(values = c("red", "blue"), name = "Sex") + 
  labs(title = "Sum of Amount Granted by Sex", x = "Sex", y = "Sum of Amount Granted") 

#Barplot to visualise the sum of the amount granted by senatorial zones
ggplot(df, aes(x = sen_zone, y = amount_granted, fill = sen_zone)) +
  geom_bar(stat = "summary", fun = "sum", position = "dodge") +
  scale_fill_manual(values = c("red", "blue", "yellow"), name = "Sen_zone") + 
  labs(title = "Sum of Amount Granted by Senatorial Zone", x = "Senatorial Zones", y = "Sum of Amount Granted") 

#density plot to understand better the distribution of the amount granted per sex
ggplot(df, aes(x = amount_granted, fill = Sex)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(values = c("red", "blue"), name = "Sex") + 
  labs(title = "Density Plot: Total Amount Granted by Sex", x = "Total Amount Granted")

#Stacked bar plot to understand how different genders contributed to the amount granted per senatorial zones

df %>%
  count(sen_zone, Sex) %>%
  ggplot(aes(x = sen_zone, y = n, fill = Sex)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("pink", "blue"), name = "Sex") + 
  labs(title = "Stacked Bar Plot: Sex by Senatorial Zones", x = "Senatorial Zone", y = "Frequency", fill = "Sex")

#Scatterplot to understand if the amount granted and percentage repaid has a positive relationship

ggplot(df, aes(x = amount_granted, y = Percentage)) +
  geom_point(color = "blue", size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Scatterplot: Amount Granted vs Percentage paid", x = "Amount Granted", y = "Percentage paid")

cor(amount_granted,Percentage)

